Eyre HTTP Caching

### 2024-09-11 mss.tex
HTTP small caps
Post w/o eaten text

%%MANUSCRIPT
\title{Eyre HTTP Caching}
\author{\patp{palfun-foslup}, \\ Ted Blackman~\patp{rovyns-ricfer}, \\ Joe Bryan~\patp{master-morzod} \\ Tlon Corporation, Urbit Foundation, Urbit Foundation}
\date{}

@@

\section{Eyre \textsc{http} Caching}

@@

\begin{enumerate}
  \item  {\bf Static content} is fully self-contained and only changes when the data within changes; for example: a blog post without comments, or an image.
  \item  {\bf Dynamic content} changes based on the current state of an agent (or some other datum); for example: a blog post \emph{with} comments or a list of \lstinline[style=inlinecode]{%pals}.
  \item  {\bf Procedural content} is generated from the request itself. While the response for any given path may be known ahead of time, it may not be possible to enumerate all the valid paths for which we have responses; for example: a parameterized \lstinline[style=inlinecode]{/sigil.svg} endpoint, or a calculator \textsc{api}.
\end{enumerate}

@@

The ``static file from Clay'' case can be implemented using the affordances here. To avoid repetition of boilerplate patterns regarding this, we might ship a small piece of userspace infrastructure, an `/app/file-server` if you will, that is responsible for bindings of this kind.

@@

But it doesn't have to be that way. The `%save` task could simply include an `auth=?` flag alongside the `data`, indicating whether authentication is required or not. The runtime would then, where needed, check the incoming request for a valid authentication cookie, and either give the response or serve a simple 403.

...Except the runtime does not presently know how to check an incoming request for authentication. And making it do so is outside the scope of the grant. The changes here aren't too big though, and in some ways similar to the behavior outlined above. (Most likely, just Eyre telling the runtime about creation/expiry of session identifiers, and teaching the runtime how to check for its presence in any `Cookie` headers.) Considering the very-nice-to-have nature of support for private endpoints, we (read: ~palfun-foslup) may offer to implement this in the short-term, so that the grant work may make use of it.

@@

\noindent\sloppy
\emph{This appendix is a response by \patp{rovyns-ricfer} and \patp{master-morzod} on \texttt{\textasciitilde 2022.8.16}.  It particularly emphasizes the role that the bound scry namespace and remote scry play in Urbit's prioritization of referential transparency.}

>>>>

%%MANUSCRIPT
\title{Eyre HTTP Caching}
\author{Mark Staarink~\patp{palfun-foslup}, \\ Ted Blackman~\patp{rovyns-ricfer}, \\ Joe Bryan~\patp{master-morzod} \\ Tlon Corporation, Urbit Foundation, Urbit Foundation}
\date{}

@@

\section[Improving Eyre \textsc{http} Caching]{Improving Eyre HTTP Caching}

@@

\begin{enumerate}
  \item  {\bf Static content} is fully self-contained and only changes when the data within changes; for example: a blog post without comments, or an image.
  \item  {\bf Dynamic content} changes based on the current state of an agent (or some other datum); for example: a blog post \emph{with} comments or a list of pals.
  \item  {\bf Procedural content} is generated from the request itself. While the response for any given path may be known ahead of time, it may not be possible to enumerate all the valid paths for which we have responses; for example: a parameterized \lstinline[style=inlinecode]{/sigil.svg} endpoint, or a calculator \textsc{api}.
\end{enumerate}

@@

The ``static file from Clay'' case can be implemented using the affordances here. To avoid repetition of boilerplate patterns regarding this, we might ship a small piece of userspace infrastructure, an \lstinline[style=inlinecode]{/app/file-server} if you will, that is responsible for bindings of this kind.

@@

But it doesn't have to be that way. The \lstinline[style=inlinecode]{%save} task could simply include an \lstinline[style=inlinecode]{auth=?} flag alongside the \lstinline[style=inlinecode]{data}, indicating whether authentication is required or not. The runtime would then, where needed, check the incoming request for a valid authentication cookie, and either give the response or serve a simple 403.

\ldots{} except that the runtime does not presently know how to check an incoming request for authentication. And making it do so is outside the scope of the grant. The changes here aren't too big though, and in some ways similar to the behavior outlined above. (Most likely, just Eyre telling the runtime about creation/expiry of session identifiers, and teaching the runtime how to check for its presence in any \lstinline[style=inlinecode]{Cookie} headers.) Considering the very-nice-to-have nature of support for private endpoints, we (read: \patp{palfun-foslup}) may offer to implement this in the short-term, so that the grant work may make use of it.

@@

\noindent\sloppy
\emph{This document is an alternative proposal by \patp{rovyns-ricfer} and \patp{master-morzod} on \texttt{\textasciitilde 2022.8.16}.  It particularly emphasizes the role that the bound scry namespace and remote scry play in Urbit's prioritization of referential transparency.}
